NLTK: The Natural Language Toolkit

### 2002-03-20 technical/acl2002/submission.tex
- Minor revisions for "design criteria" intro - commented out "python implementation" section - Added 1st draft text for modules:fsa - Added 1st draft text for modules:visualization - Revised intro to "assignments and projects" - Put chunk parsing in a subsubsection* - Commented out explanation of what code they added for chunkparser - Added 1st draft text for class demonstrations
svn/trunk@514

% To do:
%   - FSA module description
%   - visualization modules description
%   - in-class demonstrations section
%   - Get rid of Python implementation section, or at least shorten it
%   - General clean-up

@@

Several criteria must be considered in the design and implementation
of the toolkit.  The design criteria are divided into primary and
secondary criteria, and within those sections are listed in the order
of their importance.  A set of explicit non-requirements is also
given.  The toolkit is not expected to satisfy any of these
properties.

@@

\section{Python Implementation}

The structure of the Python language directly affects the design and
implementation of the toolkit.  This section discusses how the design
of the toolkit is affected by the Python language.

% [EL] This is the only section that I think we might want to keep
% from python impl.  But I'm not sure we even need to keep this one.
\noindent\textbf{Interfaces.}
Python does not directly implement interfaces.  The toolkit
therefore implements interfaces as simple classes, all of whose methods 
raise \texttt{AssertionError} exceptions.  A description of the
functionality provided by the interface is contained in the
class's documentation string.  Each method's documentation string
specifies the behavior of that method.
In order to implement an interface, a class should include that
interface as one of its bases, and should override every
(non-optional) method defined by the interface.
Classes that define interfaces are named with a trailing ``I,'' such
as \texttt{TokenizerI} or \texttt{EventI}.

% [EL] Mention that Python doesn't provide native type checking?  Or
% ditch this section? :)
\noindent\textbf{Typing.}
Type checking is an important safety device, allowing programmers to
quickly locate errors in their code.  This is especially important for
novice programmers, who have less experience with debugging.  Every
function and method defined by the toolkit will check the types of its
its arguments.  Given the impact on performance, the level of type checking
will be adjustable.

%\noindent\textbf{Advanced Language Features.}
%The use of advanced language features and programming styles could
%potentially make the toolkit much more difficult to use or to
%understand.  The fewer language features that a student must learn in
%order to use the toolkit, the faster they can start developing NLP
%systems.  The language features used by
%the toolkit are summarized by figure \ref{fig:feature}.

%\begin{figure}
%\begin{centering}
%\begin{tabular}{|l|l|}
%\hline
%\noindent\textbf{Language Feature} & \textbf{Use} \\
%\hline
%Classes & used \\
%Exceptions & error conditions only \\
%Operator Overloading & used \\
%Python 2.0 Features & used \\
%Python 2.1 Features & not used \\
%Function Arguments & minimal use \\
%Lambda Functions & minimal use \\
%Mapping and Filtering & not used \\
%List Comprehensions & used \\
%Tkinter & used \\
%\hline
%\end{tabular}\\
%\end{centering}
%\caption{Advanced language features used by NLTK.}
%\label{fig:feature}
%\end{figure}

\section{Modules}
% [EL]: I think that this section works better as a textual
% description, rather than a long list.  Also, we're going to be short
% on space.. :)
% What order should these be in?  Put more impressive stuff at the
% beginning and end?  Maybe
% parsing;tagging;visualization;classification?  or swap viz and
% classification?

@@

The remaining modules define data structures and interfaces for
performing specific NLP tasks.

@@

\texttt{fsa}.  Planning to add HMMs??

@@

\texttt{draw.tree}
\texttt{draw.tree\_edit}
\texttt{draw.plot\_graph}
\texttt{draw.chart}
\texttt{draw.fsa}

@@

%NLTK currently includes the following modules.

%\subsection*{Basics}

%\begin{description}
%\item[token:] Basic classes for processing individual
%      elements of text, such as words or sentences. 
%\item[tree:] Classes for representing tree structures
%      over text (such as syntax trees and morphological trees).
%\item[probability:] Classes that encode frequency
%      distributions and probability distributions.
%\end{description}

%\subsection*{Tagging}

%\begin{description}
%\item[tagger:] A standard interface to tag each token of
%      a text with supplementary information, such as its part of
%      speech; and several implementations of that interface.
%\end{description}

%\subsection*{Parsing}

%\begin{description}
%\item[parser:] A standard interface to produce trees
%      representing the structure of texts.
%\item[chartparser:] A flexible parser implementation
%      that uses a \emph{chart} to record hypotheses about
%      syntactic constituents.
%\item[srparser\_template:] A partial implementation of a
%      shift-reduce parser; completing the implementation was a
%      student exercise.
%\item[chunkparser:] A standard interface for robust
%      parsers used to identify non-overlapping linguistic groups
%      (such as noun phrases) in unrestricted text.
%\item[rechunkparser:] A regular-expression based
%      implementation of the chunk parser interface.
%\end{description}

%\subsection*{Text Classification}

%\begin{description}
%\item[classifier:] A standard interface for classifying
%      texts into categories.
%\item[classifier.feature:] A standard way of encoding
%      the information used to make classification decisions.
%\item[classifier.naivebayes:] A text classifier
%      implementation based on the Naive Bayes assumption.
%\item[classifier.maxent:] An implementation of the
%      maximum entropy model for text classification; and
%      implementations of the GIS and IIS algorithms for training
%      the classifier.
%\item[classifier.featureselection:] A standard interface
%      for choosing which features are relevant for making
%      classification decisions.
%\end{description}

%\subsection*{Visualization}

%\begin{description}
%\item[draw.tree:] A graphical representation for tree
%      structures, such as syntax trees and morphological trees. 
%\item[draw.tree\_edit:] A graphical interface used to
%      build and modify tree structures.
%\item[draw.plot\_graph:] A graphical tool to graph
%      arbitrary functions.
%\item[draw.chart:] An interactive graphical tool used to
%      experiment with chart parsers.
%\end{description}

@@

NLTK can be used for simple and advanced projects, as well as
everything in between.  In the simplest assignments, students tweak an
existing module to alter its behavior.  The variety of existing
modules makes it easy to create simple assignments.  A more
challenging task is to develop a new module.  Here, NLTK provides some
useful starting points: predefined interfaces and data structures, and
existing modules that implement the same interface.  Advanced projects
typically involve the development of entirely new functionality for a
previously unsupported NLP task, or the development of a complete
system out of existing and new modules.

Here we illustrate the use of NLTK in an assignment of moderate
difficulty, for the case of chunk parsing.  This is a kind of parsing
where we only identify the main constituents of a phrase, a technique
with diverse applications.  The NLTK \texttt{chunk} module offers
a variety of rule formats.  For example,
\texttt{ChunkRule('<NN.*>')} builds chunks over consecutive noun tags;
\texttt{ChinkRule('<VB.>')} excises verbs from existing chunks;
\texttt{SplitRule('<NN>', '<DT>')} splits into two pieces any existing
chunk that contains a noun followed by a determiner; and
\texttt{MergeRule('<JJ>', '<JJ>')} combines two adjacent chunks
where the first chunk ends and the second chunk starts with \texttt{JJ}.
The goal of the assignment is to come up with a cascade of these rules,
and to create a set of chunks that are as similar as possible to the test set.

@@

% [EL] I'm not sure this is necessary:
The main program which surrounds the student's work is moderately
simple as well.  The line replaced by students is underlined.
% [EL] I find the undeline hard to read; change to italics or bold or
% vbar in the margin?  Or maybe a \Rightarrow of some sort in the margin?
\begin{sv}
from nltk.token import *
from nltk.chunkparser import *
from nltk.rechunkparser import *
import os

path = "/cis530/data/wsj/tagged/"

sent_tok = RETokenizer('(=======+)?{\textbackslash}{\textbackslash}n{\textbackslash}{\textbackslash}n+', 0)
sentences = []
files = os.listdir(path)
for file in files:
    print 'Loading \%s...' \% file
    text = open(path+file).read()
    sentences += sent_tok.tokenize(text, source=file, unit='s')

chunkscore = ChunkScore()

\underline{cascade = [ ??? ]}
chunkparser = REChunkParser(cascade)

ctt = ChunkedTaggedTokenizer()
for s in sentences:
    correct = ctt.tokenize(s.type(), source=s.loc())
    unchunked = unchunk(correct)
    guess = chunkparser.parse(unchunked)
    chunkscore.score(correct, guess)

print chunkscore
\end{sv}

The score for each project is the precision, recall and F-measure.
Figure~\ref{contest} shows the results; students with the best two scores
were presented with prizes.

\begin{figure}
\centerline{\epsfig{figure=contest.ps,width=\linewidth}}
\caption{Precision/Recall Graph for Chunking Competition}\label{contest}
\vspace*{2ex}\hrule
\end{figure}

@@

[chart parsing example, including screenshot]

>>>>

% To do:
%   - FSA module description
%   - visualization modules description
%   - in-class demonstrations section
%   - Get rid of Python implementation section, or at least shorten it
%   - General clean-up
%   - ispell

@@

Several criteria were considered in the design and implementation of
the toolkit.  The design criteria are divided into primary and
secondary criteria, and within those sections are listed in the order
of their importance.  It was also important to decide what goals the
toolkit would \emph{not} attempt to accomplish; we therefore include
an explicit set of non-requirements, which the toolkit is not expected
to satisfy.

@@

%\section{Python Implementation}
%
%The structure of the Python language directly affects the design and
%implementation of the toolkit.  This section discusses how the design
%of the toolkit is affected by the Python language.
%
%% [EL] This is the only section that I think we might want to keep
%% from python impl.  But I'm not sure we even need to keep this one.
%\noindent\textbf{Interfaces.}
%Python does not directly implement interfaces.  The toolkit
%therefore implements interfaces as simple classes, all of whose methods 
%raise \texttt{AssertionError} exceptions.  A description of the
%functionality provided by the interface is contained in the
%class's documentation string.  Each method's documentation string
%specifies the behavior of that method.
%In order to implement an interface, a class should include that
%interface as one of its bases, and should override every
%(non-optional) method defined by the interface.
%Classes that define interfaces are named with a trailing ``I,'' such
%as \texttt{TokenizerI} or \texttt{EventI}.
%
%% [EL] Mention that Python doesn't provide native type checking?  Or
%% ditch this section? :)
%\noindent\textbf{Typing.}
%Type checking is an important safety device, allowing programmers to
%quickly locate errors in their code.  This is especially important for
%novice programmers, who have less experience with debugging.  Every
%function and method defined by the toolkit will check the types of its
%its arguments.  Given the impact on performance, the level of type checking
%will be adjustable.
%
%\noindent\textbf{Advanced Language Features.}
%The use of advanced language features and programming styles could
%potentially make the toolkit much more difficult to use or to
%understand.  The fewer language features that a student must learn in
%order to use the toolkit, the faster they can start developing NLP
%systems.  The language features used by
%the toolkit are summarized by figure \ref{fig:feature}.
%
%\begin{figure}
%\begin{centering}
%\begin{tabular}{|l|l|}
%\hline
%\noindent\textbf{Language Feature} & \textbf{Use} \\
%\hline
%Classes & used \\
%Exceptions & error conditions only \\
%Operator Overloading & used \\
%Python 2.0 Features & used \\
%Python 2.1 Features & not used \\
%Function Arguments & minimal use \\
%Lambda Functions & minimal use \\
%Mapping and Filtering & not used \\
%List Comprehensions & used \\
%Tkinter & used \\
%\hline
%\end{tabular}\\
%\end{centering}
%\caption{Advanced language features used by NLTK.}
%\label{fig:feature}
%\end{figure}

\section{Modules}
% [EL]: I think that this section works better as a textual
% description, rather than a long list.  Also, we're going to be short
% on space.. :)
% What order should these be in?  Put more impressive stuff at the
% beginning and end?

@@

The remaining modules define data structures and interfaces for
performing specific NLP tasks.  This list of modules will grow over
time, as we add new tasks and algorithms to the toolkit.

@@

% Steven: do you want to add anything here?  Say anything about HMMS? 
% Will you have implemented more interfaces by the time of ACL?
The \texttt{fsa} module defines a data type for encoding finite state
automata; and an interface for creating automata from
regular expressions.

@@

A set of visualization modules define graphical interfaces for viewing
and editing data structures; and graphical tools for experimenting
with tasks.  The \texttt{draw.tree} module provides a
simple graphical interface for displaying tree structures; and the
\texttt{draw.tree\_edit} module provides an interface for building and
modifying tree structures.  The \texttt{draw.plot\_graph} module can be
used to graph mathematical functions.  The \texttt{draw.fsa} module
provides a graphical tool for displaying and simulating finite state
automata.  The \texttt{draw.chart} module provides an interactive
graphical tool for experimenting with chart parsers.

@@

%% Old content for this section:
%NLTK currently includes the following modules.
%
%\subsection*{Basics}
%
%\begin{description}
%\item[token:] Basic classes for processing individual
%      elements of text, such as words or sentences. 
%\item[tree:] Classes for representing tree structures
%      over text (such as syntax trees and morphological trees).
%\item[probability:] Classes that encode frequency
%      distributions and probability distributions.
%\end{description}
%
%\subsection*{Tagging}
%
%\begin{description}
%\item[tagger:] A standard interface to tag each token of
%      a text with supplementary information, such as its part of
%      speech; and several implementations of that interface.
%\end{description}
%
%\subsection*{Parsing}
%
%\begin{description}
%\item[parser:] A standard interface to produce trees
%      representing the structure of texts.
%\item[chartparser:] A flexible parser implementation
%      that uses a \emph{chart} to record hypotheses about
%      syntactic constituents.
%\item[srparser\_template:] A partial implementation of a
%      shift-reduce parser; completing the implementation was a
%      student exercise.
%\item[chunkparser:] A standard interface for robust
%      parsers used to identify non-overlapping linguistic groups
%      (such as noun phrases) in unrestricted text.
%\item[rechunkparser:] A regular-expression based
%      implementation of the chunk parser interface.
%\end{description}
%
%\subsection*{Text Classification}
%
%\begin{description}
%\item[classifier:] A standard interface for classifying
%      texts into categories.
%\item[classifier.feature:] A standard way of encoding
%      the information used to make classification decisions.
%\item[classifier.naivebayes:] A text classifier
%      implementation based on the Naive Bayes assumption.
%\item[classifier.maxent:] An implementation of the
%      maximum entropy model for text classification; and
%      implementations of the GIS and IIS algorithms for training
%      the classifier.
%\item[classifier.featureselection:] A standard interface
%      for choosing which features are relevant for making
%      classification decisions.
%\end{description}
%
%\subsection*{Visualization}
%
%\begin{description}
%\item[draw.tree:] A graphical representation for tree
%      structures, such as syntax trees and morphological trees. 
%\item[draw.tree\_edit:] A graphical interface used to
%      build and modify tree structures.
%\item[draw.plot\_graph:] A graphical tool to graph
%      arbitrary functions.
%\item[draw.chart:] An interactive graphical tool used to
%      experiment with chart parsers.
%\end{description}

@@

NLTK can be used to create student assignments of varying difficulty
and scope.  
% Use a module
In the simplest assignments, students experiment with an existing
module.  The wide variety of existing modules provide many opportunies
for creating these simple assignments.
% Edit/extend a module.
Once students become more familiar with the toolkit, they can be asked
to make minor changes or extensions to an existing module.
% Develop a new module.
A more challenging task is to develop a new module.  Here, NLTK
provides some useful starting points: predefined interfaces and data
structures, and existing modules that implement the same interface.
% Advanced projects..
Advanced projects typically involve the development of entirely new
functionality for a previously unsupported NLP task, or the
development of a complete system out of existing and new modules.

\subsubsection*{Example: Chunk Parsing}

% Akward
Here we illustrate the use of NLTK in an assignment of moderate
difficulty, for the case of chunk parsing.  This is a kind of parsing
where we only identify the main constituents of a phrase, a technique
with diverse applications.  The NLTK \texttt{chunk} module offers
a variety of rule formats.  For example,
\texttt{ChunkRule('<NN.*>')} builds chunks over consecutive noun tags;
\texttt{ChinkRule('<VB.>')} excises verbs from existing chunks;
\texttt{SplitRule('<NN>', '<DT>')} splits into two pieces any existing
chunk that contains a noun followed by a determiner; and
\texttt{MergeRule('<JJ>', '<JJ>')} combines two adjacent chunks
where the first chunk ends and the second chunk starts with \texttt{JJ}.
The goal of the assignment is to come up with a cascade of these rules,
and to create a set of chunks that are as similar as possible to the test set.

@@

%% [EL] I'm not sure this is necessary:
%The main program which surrounds the student's work is moderately
%simple as well.  The line replaced by students is underlined.
%% [EL] I find the undeline hard to read; change to italics or bold or
%% vbar in the margin?  Or maybe a \Rightarrow of some sort in the margin?
%\begin{sv}
%from nltk.token import *
%from nltk.chunkparser import *
%from nltk.rechunkparser import *
%import os
%
%path = "/cis530/data/wsj/tagged/"
%
%sent_tok = RETokenizer('(=======+)?{\textbackslash}{\textbackslash}n{\textbackslash}{\textbackslash}n+', 0)
%sentences = []
%files = os.listdir(path)
%for file in files:
%    print 'Loading \%s...' \% file
%    text = open(path+file).read()
%    sentences += sent_tok.tokenize(text, source=file, unit='s')
%
%chunkscore = ChunkScore()
%
%\underline{cascade = [ ??? ]}
%chunkparser = REChunkParser(cascade)
%
%ctt = ChunkedTaggedTokenizer()
%for s in sentences:
%    correct = ctt.tokenize(s.type(), source=s.loc())
%    unchunked = unchunk(correct)
%    guess = chunkparser.parse(unchunked)
%    chunkscore.score(correct, guess)
% 
%print chunkscore
%\end{sv}

The score for each project is the precision, recall and F-measure.
Figure~\ref{fig:contest} shows the results; students with the best two
scores were presented with prizes.

\begin{figure}
\centerline{\epsfig{figure=contest.ps,width=\linewidth}}
\caption{Precision/Recall Graph for Chunking Competition}\label{fig:contest}
\vspace*{2ex}\hrule
\end{figure}

@@

NLTK provides graphical tools that can be used to give in-class
demonstrations, which help to explain basic NLP concepts and
algorithms.  These interactive tools can be used to display relevant
data structures, and to show step-by-step executions of algorithms.
Both data structures and control flow can be easily modified during
the demonstration, in response questions from the class.

Since these graphical tools are included with the toolkit, they can
also be used by students.  This allows students to experiment with the
algorithms that they have seen presented in class.

\subsubsection*{Example: The Chart Parsing Tool}

The chart parsing tool can be used to explain the basic concepts
behind chart parsing, and to show how the algorithm works.  Chart
parsing is a flexible parsing algorithm that uses a data structure
called a \emph{chart} to record hypotheses about syntactic
constituents.  Each hypothesis is represented by a single \emph{edge}
on the chart.  A set of \emph{rules} determine when new edges can be
added to the chart.  This set of rules controls the overall behavior
of the parser (e.g., whether it parses top-down or bottom-up).

The chart parsing tool demonstrates the process of parsing a single
sentence, with a given grammar and lexicon.  Its display is divided
into three sections: the bottom section displays the chart; the middle
section displays the sentence; and the top section displays the
partial syntax tree corresponding to the selected edge.  Buttons along
the bottom of the window are used to control the execution of the
algorithm.  The main display window for the chart parsing tool is
shown in Figure~\ref{fig:chartparse}.

This tool can be used to explain several different aspects of chart
parsing.  First, it can be used to explain the basic chart data
structure, and to show how edges can represent hypotheses about
syntactic constituents.  It can then be used to demonstrate and
explain the individual rules that the chart parsers uses to create new
edges.  Finally, it can be used to show how these individual rules
combine to find a complete parse for the sentence.

% Is ``user'' a good word here?  ``lecturer''?
To reduce the overhead of setting up demonstrations during lecture,
the user can define a list of preset charts.  The tool can then be
reset to any one of these charts at any time.

The chart parsing tool allows for flexible control of the parsing
algorithm.  At each step of the algorithm, the user can select which
rule or strategy they wish to apply.  This allows the user to
experiment with mixing different strategies (e.g., top-down and
bottom-up).  The user can exercise fine-grained control over the
algorithm by selecting which edge they wish to apply a rule to.  This
flexibility allows lecturers to use the tool to respond to a wide
variety of questions; and allows students to experiment with different
variations on the chart parsing algorithm.

\begin{figure}
%\centerline{\epsfig{figure=chart.ps,width=\linewidth}}
\caption{Chart Parsing Tool}\label{fig:chartparse}
\vspace*{2ex}\hrule
\end{figure}
